---
title: 'STAT 420: Homework 02'
author: "Spring 2020, Jung Hyun Hwang"
date: 'February 9 '
output:
  html_document:
    theme: readable
    toc: yes
  pdf_document:
    toc: yes
---

## Exercise 1 (Writing Simple Functions)

For each of the following parts, use the following vectors:

```{r}
a = 1:10
b = 10:1
c = rep(1, times = 10)
d = 2 ^ (1:10)
```

**(a)** Write a function called `sum_of_squares`.

- Arguments:
    - A vector of numeric data `x`.
- Output:
    - The sum of the squares of the elements of the vector. $\sum_{i = 1}^n x_i^2$
    
Provide your function, as well as the result of running the following code:

```{r, eval = FALSE}
sum_of_squares(x = a)
sum_of_squares(x = c(c, d))
```

```{r}
sum_of_squares = function(x) {
  result = sum(x^2)
  result
}

sum_of_squares(a)
sum_of_squares(c(c, d))
```



**(b)** Write a function called `sum_of_power`.

- Arguments:
    - A vector of numeric data `x`.
    - `p` which should have the default value of `2`.
- Output:
    - $\sum_{i = 1}^n x_i^{p}$
    
Provide your function, as well as the result of running the following code:

```{r, eval = FALSE}
sum_of_power(x = a)
sum_of_power(x = a, p = 3)
sum_of_power(x = a, p = a)
sum_of_power(x = a, p = c(1, 2))
```

```{r}
sum_of_power = function(x, power = 2) {
  result = sum(x^power)
  result
}
  
sum_of_power(x = a)
sum_of_power(x = a, p = 3)
sum_of_power(x = a, p = a)
sum_of_power(x = a, p = c(1, 2))
```


**(c)** Write a function called `rms_diff`.

- Arguments:
    - A vector of numeric data `x`.
    - A vector of numeric data `y`.
- Output:
    - $\sqrt{\frac{1}{n}\sum_{i = 1}^n (x_i - y_i)^{2}}$
    
Provide your function, as well as the result of running the following code:

```{r, eval = FALSE}
rms_diff(x = a, y = b)
rms_diff(x = d, y = c)
rms_diff(x = d, y = 1)
rms_diff(x = a, y = 0) ^ 2 * length(a)
```

```{r}
rms_diff = function(x, y) {
  result = -sqrt(sum((x - y)^2) / length(x))
  result
}

rms_diff(x = a, y = b)
rms_diff(x = d, y = c)
rms_diff(x = d, y = 1)
rms_diff(x = a, y = 0) ^ 2 * length(a)
```

## Exercise 2 (Plotting, Testing)

For this exercise we will use [the data that is stored in `intelligence.csv`](intelligence.csv) which records IQs of a random sample of residents of Pawnee and Eagleton, Indiana.

**(a)** Load the data from `intelligence.csv` into a variable in `R` called `intelligence`. Show the code used to do this.
```{r}
library(readr)
intelligence <- read_csv("intelligence.csv")
```


**(b)** Create a side-by-side boxplot that compares the IQs across the two towns. Be sure to give the plot a title and label the axes appropriately.
```{r}
boxplot(iq ~ town, data = intelligence,
        ylab = 'IQ',
        xlab = 'town',
        main = 'IQs vs town',
        col = 'brown')
```


**(c)** Are people from Eagleton smarter than people from Pawnee? Perform an appropriate statistical test using the given sample data. That is, test $H_0: \mu_E = \mu_P$ vs $H_1: \mu_E > \mu_P$, where

- $\mu_E$ is the mean IQ of a resident of Eagleton.
- $\mu_P$ is the mean IQ of a resident of Pawnee.

Explicitly state the p-value of the test and the resulting statistical decision at a significance level $\alpha = 0.10$. Interpret the results in the context of the problem.
```{r}
Eagleton = intelligence$iq[intelligence$town == 'eagleton']
Pawnee = intelligence$iq[intelligence$town == 'pawnee']

result = t.test(x = Eagleton, mu = mean(Pawnee), alternative = c('greater'), conf.level = 0.9)
result
result$p.value
```

Since p-value is greater than alpha value, it fails to reject. It is not possible to figure out people from Eagleton are smarter than people from Pawnee.

**(d)** Repeat **(c)** using a two-sided alternative hypothesis. What changes?
```{r}
Eagleton = intelligence$iq[intelligence$town == 'eagleton']
Pawnee = intelligence$iq[intelligence$town == 'pawnee']

result = t.test(Eagleton, Pawnee, alternative = 'two.sided')
result
result$p.value
```

p-value increased when it is compared to c. Since p-value is still greater than alpha value, it fails to reject. It is still impossible to tell people from Eagleton are smarter than people from Pawnee.

## Exercise 3 (Writing More Functions)

In this exercise we will write our own functions related to performing a one-sample $t$ test. That is $H_{0}: \mu = \mu_{0}$ versus $H_{1}: \mu \neq \mu_{0},$ where $\mu_{0}$ is the hypothesized value of $\mu$.

Throughout this exercise you may **not** use the `t.test()` function inside your functions. You may use it to check your work separately, but no such double-checks should appear in your final report.

Some built in `R` functions that may be useful to you when writing your functions include: `c()`, `ifelse()`, `mean()`, `sd()`, `abs()`, `length()`, `sqrt()`, and `pt()`.

**(a)** Write a function called `do_t_test` which takes two inputs:

- `x`: A vector which stores observations.
- `mu`: The hypothesized value of $\mu$ which defaults to `0`.

The function should output:

- The value of the test statistic, $t$.
- The p-value of the test. The function only needs to be able to handle a two-sided alternative.

In order to output both, consider using `c(t, pval)` as the last line of your function, and store those two values elsewhere in the body of your function.
```{r}
do_t_test = function(x, mu = 0) {
  n = length(x)
  x_bar = mean(x)
  sample_sd = sd(x)
  t = abs(sqrt(n) * (x_bar - mu) / sample_sd)
  pval = 2 * pt((-1) * t, df = (n - 1))
  c(t, pval)
}
```


**(b)** Write a function called `make_decision` which takes two inputs:

- `pval`: The p-value of a test.
- `alpha`: The significance level of a test. Set a default value of `0.05`.

The function should output `"Reject!"` or `"Fail to Reject."` based on the comparison of `pval` to `alpha`.
```{r}
make_decision = function(pval, alpha = 0.05) {
  ifelse(pval < alpha, 'Reject', 'Fail to Reject')
}
```


**(c)** Now we will test the quality of your functions from parts **(a)** and **(b)**. Run the following code:

```{r, eval = FALSE}
set.seed(114)
y = rnorm(25, 1.4, 1)
pval = do_t_test(y, mu = 2)[2]
pval
make_decision(pval, alpha = 0.10)
```
```{r}
set.seed(114)
y = rnorm(25, 1.4, 1)
pval = do_t_test(y, mu = 2)[2]
pval
make_decision(pval, alpha = 0.10)
```

If your `do_t_test()` and `make_decision()` functions are correct, you should obtain a decision of `"Fail to Reject."` You will also be evaluated on whether the numeric p-value you obtain is correct.





## Exercise 4 (CLT Simulation)

For this exercise we will simulate from the exponential distribution. If a random variable $X$ has an exponential distribution with rate parameter $\lambda$, the pdf of $X$ can be written

\[
f(x;\lambda) = \lambda e ^ {-\lambda x}
\]

for $x \geq 0$.

Also recall,

\[
\begin{aligned}
\mu &= E[X] = \frac{1}{\lambda} \\
\sigma^ 2 &= Var[X] = \frac{1}{\lambda ^ 2}
\end{aligned}
\]

**(a)** This exercise relies heavily on generating random observations. To make this reproducible we will set a seed for the randomization. Alter the following code to make `birthday` store your birthday in the format: `yyyymmdd`. For example, [William Gosset](https://en.wikipedia.org/wiki/William_Sealy_Gosset), better known as *Student*, was born on June 13, 1876, so he would use:

```{r}
birthday = 19960729
set.seed(birthday)
```

**(b)** Simulate 10000 samples of size **5** from an exponential distribution with $\lambda = 2$. Store the mean of each sample in a vector. Plot a histogram of these sample means. (Be sure to give it a title, and label the axes appropriately.) Based on the histogram, do you think the central limit theorem applies here?
```{r}
N1 = 1000
n1 = 5
lambda = 2
x1_bar = {}
for(i in 1:N1) {
  y = rexp(n1, lambda)
  x1_bar[i] = mean(y)
}
hist(x1_bar,
     xlab = 'sample mean',
     ylab = 'frequencey,',
     main = 'Histogram',
     col = 'brown')
```

It is quite hard to say the central limit theorem applies here based on the histogram. It is probably because n is not big enough.

**(c)** Simulate 10000 samples of size **100** from an exponential distribution with $\lambda = 2$. Store the mean of each sample in a vector. Plot a histogram of these sample means. (Be sure to give it a title, and label the axes appropriately.) Based on the histogram, do you think the central limit theorem applies here?
```{r}
N2 = 1000
n2 = 100
lambda = 2
x2_bar = {}
for(i in 1:N2) {
  y2 = rexp(n2, lambda)
  x2_bar[i] = mean(y2)
}
hist(x2_bar,
     xlab = 'sample mean',
     ylab = 'frequencey,',
     main = 'Histogram',
     col = 'brown')
```

Yes, the central limit theorem applies here based on the histogram. It has much greater n value than previous simulation.

**(d)** We just repeated ourselves, so that means we probably should be writing a function. Write a function called `sim_xbars_exp` which takes three inputs:

- The number of samples to simulate.
- The sample size.
- The rate parameter of an exponential distribution.

The function should output a vector of sample means which are the result of sampling from an exponential distribution as specified by the inputs.

Use your function to simulate 25000 samples of size **50** from an exponential distribution with $\lambda = 3$. Store the mean of each sample in a vector. Plot a histogram of these sample means. (Be sure to give it a title, and label the axes appropriately.)
```{r}
sim_xbars_exp = function(N, n, lambda) {
  set.seed(19960729)
  x_bar = {}
  for(i in 1:N) {
    y3 = rexp(n, lambda)
    x_bar[i] = mean(y3)
  }
  x_bar
}
h = sim_xbars_exp(25000, 50, 3)
hist(h,
     ylab = 'Frequencey',
     xlab = 'Sample mean',
     main = 'Histogram',
     col = 'brown')
```


## Exercise 5 (More Simulation)

Let $X$ follow an exponential distribution with rate parameter $\lambda_X = 2$. Let $Y$ follow a Poisson distribution with rate parameter $\lambda_Y = 3$.

We write $sd(X)$ for the true standard deviation of $X$ and $m(Y)$ for the true median of $Y$.

Let $s_x$ be the sample standard deviation of $X$ which is an estimate of $sd(X)$. Also let $m_y$ be the sample median which is an estimate of $m(Y)$.

Suppose we take samples of size $n_x = 10$ from $X$ and take samples of size $n_y = 5$. Consider the statistic

\[
\frac{s_x}{m_y}.
\]

What is the (sampling) distribution of $\frac{s_x}{m_y}$? We could ask a statistician who specializes in theory. Instead of using mathematics, simulate $\frac{s_x}{m_y}$ 5000 times and store the results. Plot a histogram of the observed values of $\frac{s_x}{m_y}$. Comment on the shape of the histogram and empirical distribution of $\frac{s_x}{m_y}$. Before running your code, set the same seed used for the previous exercise. For full credit, do **not** use a `for` loop.
```{r}
set.seed(19960729)
lambda_x = 2
lambda_y = 3
n_x = 10
n_y = 5
x = rexp(n_x, lambda_x)
y = rpois(n_y, lambda_y)
t = sd(x) / median(y)
N = 5000

m_xx = matrix(rexp(n_x * 5000, lambda_x), nrow = 5000, byrow = T)
s_x = apply(m_xx, 1, sd)

m_yy = matrix(rpois(n_y * 5000, lambda_y), nrow = 5000, byrow = T)
m_y = apply(m_yy, 1, median)

t = s_x / m_y
hist(t,
     ylab = 'Frequency',
     xlab = 't value',
     main = 'Histogram',
     col = 'brown'
     )
```

The probabilith distrivution of a statistic is called sampling distribution of that statistic. This distribution is extremely right skewed. Frequency tends to be dramatically decrease as t-value passes 0.2. Majority of data falls between t-value 0.0 and 0.2.